---
title: "lab-07-simpsons.Rmd"
author: "Fatimah-Albrahim "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham

```
    
Your answer:
The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)

```

Your answer;
there are 1314 observations . as we know every row is an observation

3.  

```{r number_variables}
names(Whickham)

```

Your answer:  
there are 3, "outcome", "smoker",  "age"

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)

```

Your answer: 
using the 'unique()' funtion on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +

 geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +

 geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age )) +
 geom_histogram(binwidth = 10 , color = "black")

```


4. i expect the health will be worser and may be the person will be died after while , if he keeping smoke .

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
 count(smoker, outcome) %>%
group_by(smoker) %>%
  mutate(prop_outcome = n / sum(n)) %>%
  filter(outcome=="dead")

ggplot(Whickham, aes(x=smoker, fill=outcome)) + 
  geom_bar(position = "fill") + labs(y="proprtion") + theme_dark()


```

6.

```{r age_cat}

Whickham <- Whickham%>% mutate (age_cat = case_when (age <= 44 ~ "18-44", age > 44. & age <= 64 ~ "45-64", age > 64 ~ "65+"))

```
 
7. 

```{r plot_smokeVSincome_3}

ggplot(Whickham, aes(x=smoker, fill=outcome)) + geom_bar(position = "fill") + 
  labs(y="Proportion") + facet_grid(. ~ age_cat) + theme_dark()
```

Knit, commit, and push to github.