---
title: "lab-07-simpsons.Rmd"
author: "Fatimah-Albrahim "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham

```
    
Your answer:
The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)

```

Your answer;
there are 1314 observations . as we know every row is an observation

3.  

```{r number_variables}
names(Whickham)

```

Your answer:  
there are 3, "outcome", "smoker",  "age"

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)

```

Your answer: 
using the 'unique()' funtion on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +

 geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +

 geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age )) +
 geom_histogram(binwidth = 10 , color = "black")

```


4. 

```{r plot_smokeVSincome_1}

```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
```

6.

```{r age_cat}

```

7. 

```{r plot_smokeVSincome_3}

```

Knit, commit, and push to github.